The equal distribution of replicated genetic materials packed within chromosomes into daughter cells is the hallmark of mitosis. Failure to maintain such genetic stability is detrimental and often associated with cancers. This important role is carried out by the spindle and safeguarded by signaling pathways including the spindle assembly checkpoint (SAC). The spindle attaches to chromosomes through kinetochores to prepare for distribution. The SAC monitors the completeness of attachment at all kinetochores. How the SAC blocks the progression of mitosis in the presence of a single unattached kinetochore and releases the block promptly once all attachment is secured remains unclear. Here, I summarize our discovery that the SAC, like many other biochemical switches, exploits cooperativity to achieve such sensitivity. Furthermore, I look into potential molecular mechanisms of this cooperativity and discuss future studies that may finally solve the puzzle of cooperativity in the SAC.

% The biochemical pathway of SAC has been majorly elucidated
% sensitivity ~ responsiveness
% Describe dissertation and any additional information they wish to bring to the attention of the selection committee. The statement should address the importance of the student’s work in the beginning two or three sentences. The statement should include the theoretical framework of the dissertation, its specific aims, methodologies , originality, and the significance and contribution of the project to the field. 
% Reviewers will NOT necessarily be familiar with the technical vocabulary of a specific field.
% Quality of the nominee’s personal statement, including the research plan, progress on the project, scholarly significance of the dissertation research, and, if Barbour eligible, contribution towards the student’s home country.
% Results of early research or scholarship in graduate studies, as represented; e.g., through article publication, conference presentation, or other scholarly products such as websites, performances, exhibitions, etc.
% Excellence in graduate teaching and/or pedagogy/mentoring/professional leadership and service.

During cell division, the unequal distribution of replicated genetic materials packed within chromosomes results in abnormal numbers of chromosomes (aneuploidy) in daughter cells, which is often associated with cancers. Multiple key surveillance pathways regulate the cell cycle to ensure genetic stability. One of these pathways is called the spindle assembly checkpoint (SAC), which my thesis research revolves around. A complete understanding of how the SAC safeguards against aneuploidy will teach us how to prevent or induce its failure, which may inspire novel methods to prevent or to treat cancers.

Through decades of study, the field has mostly elucidated where the SAC fits into the cell cycle and the biochemical reactions associated with the SAC. To prepare for the equal distribution of chromosomes, the dividing cell forms the spindle which attaches from opposite poles of the cell to all chromosomes through adaptor structures named kinetochores. Unattached kinetochores activate the SAC through a series of biochemical reactions. First, unattached kinetochores recruit the kinase MPS1 which phosphorylates multiple consensus sites named MELT motifs on the scaffold protein KNL1. Phosphorylated MELT motifs will then recruit a series of SAC proteins and eventually generate a diffusible inhibitory signaling molecule known as the Mitotic Checkpoint Complex (MCC) which blocks the progression of cell division. This block allows time for the spindle to find the loose kinetochores. Once all attachment is secured, the SAC will inactivate and the block will be released. The spindle will then pull the two sets of chromosomes towards opposite poles, which later become parts of the two daughter cells.

Despite the deep knowledge of the participating proteins and reactions, a fundamental question about the SAC remains unclear. How does the SAC, a biochemical signaling cascade, approaches the sensitivity of a binary electronic logic gate, blocking the progression of cell division in the presence of even just a single unattached kinetochore and releasing the block promptly once all attachment is secured? Taking a cue from different well-studied biological decision-making processes, I hypothesized that cooperativity boosts the sensitivity of the SAC like in some other scenarios. Just like effective collaborations between people in the society, positive cooperativity in biochemical processes can synergize reactions and amplify the output. This synergy and amplification may be fundamental to the sensitivity of the SAC.

Here, I and my collaborators deployed cell biology (including engineered probe system, gene editing, and live-cell time-lapse imaging), biochemistry (including immunoprecipitation and in vitro reconstitution of biochemical reactions), and biophysics (including fluorescence lifetime imaging and mathematical modeling) approaches to study if and how multiple proteins and reactions cooperate to enable the sensitivity of the SAC.

To confirm if positive cooperativity plays a role in the SAC, we first need a quantitative readout of the input (how many MELT motifs are phosphorylated to activate the SAC) and the output (the duration of cell division as a common indicator of the MCC assembly rate) with sufficient statistical power. Because of the nanoscale volume of a kinetochore where KNL1 proteins reside, gauging phosphorylation of MELT motifs on KNL1 is technically difficult in live cells. To circumvent this obstacle, we engineered a cytosolic probe to ectopically activate the SAC (Chen et al., 2019). This probe utilizes chemically-induced dimerization of MPS1 and fragments of KNL1 that contain various numbers of MELT motifs. Upon the addition of the dimerization reagent, MPS1 comes into proximity with MELT motifs of KNL1 and phosphorylates them, thereby activating the SAC. The probe system is tagged with fluorescent proteins, allowing us to correlate the duration of cell division with the concentration of the SAC-activating probe at a single-cell resolution in real time by microscopy. Our dose-response analyses surprisingly revealed that under certain conditions the SAC signaling strength is stronger with a smaller number of phosphorylated MELT motifs present. This observation of "less is more" strongly indicates positive cooperativity in the SAC, which together with the limiting abundance of certain SAC protein(s) provides a perfect theory to explain the phenomenon. This theoretical framework projects a nonlinear enrichment of certain SAC proteins in the presence of a small number of unattached kinetochores under the physiological settings, which we verified experimentally. To our knowledge, this is the first published evidence supporting the existence of cooperativity in the SAC.

After confirming the cooperativity in the SAC, we move on to inquire into its potential molecular mechanisms and prospective factors that may affect it. Based on our model, we are investigating (1) the distance between MELT motifs and the presence of auxiliary motifs of KNL1, (2) the parallel recruitment pathway of an MCC component named BUBR1, and (3) a potential spatial and temporal coupling of two reactions involving two other MCC components named CDC20 and MAD2.

We started by varying the distance between two MELT motifs in our aforementioned probe system by incorporating various numbers of tandem repeats of a specified linker. Based on tested cases, variation in the linker length between two MELT motif does not seem to yield much difference in the SAC dose-response. We further tested whether auxiliary motifs of KNL1 facilitate cooperativity in the SAC. In contradiction to previous publications, we discovered that these auxiliary motifs of KNL1 sequester BUBR1 instead of contributing to the SAC. This limits the availability of BUBR1, leading to a decreased SAC response. Together with our first paper, we concluded that the presence and the number of different functional motifs on the KNL1 affect the cooperativity in the SAC.

Next, we looked into whether the parallel recruitment pathway of BUBR1 contributes to the cooperativity in the SAC. Some BUBR1 proteins bind to phosphorylated MELT motifs which also recruit other SAC proteins that make up the MCC. We hypothesized that this pool of BUBR1 comes into proximity with other MCC components, increasing their apparent concentrations and thus facilitating the MCC assembly. We knocked down the endogenous BUBR1 and rescued the cell with either the wild-type BUBR1 or a mutated BUBR1 that can only bind to phosphorylated MELT motifs exclusively. Comparing the SAC response of the wild-type BUBR1 with that of the mutated BUBR1, our preliminary data showed that the parallel recruitment pathway of BUBR1 does not contribute to the SAC. However, a more thorough experiment is needed to seal the conclusion of this argument.

Lastly, we are currently exploring whether the potential spatial and temporal coupling between the conformational change of MAD2 and the formation of CDC20-MAD2 dimer boosts the efficiency of CDC20-MAD2 dimerization (and hence the MCC assembly). It is known that mixing CDC20 and MAD2 alone in vitro results in a slow dimerization kinetic. In human cells, CDC20 and MAD2 bind to the same protein scaffold MAD1. We hypothesized that the structural flexibility of MAD1 helps to position CDC20 and MAD2 in proximity and maneuver them, thereby catalyzing the conformational change of MAD2 and coupling this process with the CDC20-MAD2 dimerization. Our preliminary data showed that disrupting the structural flexibility of MAD1 impairs the efficiency of the MCC assembly. We are actively working to reconstitute these reactions and validate this finding in vitro.

% The SAC monitors the completeness of the attachment process and activates to block the progession of somatic cell division (mitosis) if it detects unattached kinetochores.
% Over the course of attachment, the number of unattached kinetochores decreases from 92 (when all kinetochores are unattached) to 0 (when the attachment is complete) in human cells.
% However, doing so turned out to disrupt the genetic stability of the construct. Hence, we were only able to test a limited array of linker lengths. 
% assembled from BUBR1, BUB3, CDC20, and MAD2 in human cells
%  for the non-monotonic response between the duration of mitosis and the concentration of the SAC-activating probe
% Needs better introduction of conformational change of MAD2.
% It may also spark innovations in biological circuit design and protein engineering.
% Just like one will not completely understand how a machine works just by knowing all the gears involved, we cannot claim to have completely comprehended how the SAC works by just knowing all associated biochemical reactions individually. The understanding of the entire picture is hindered by multiple difficulties. First, the biochemical reactions that lead to MCC formation are spatially localized within the nanoscale volume of a kinetochore. Second, the concentrations of many SAC proteins are tightly regulated and may be influenced by each other. Third, that the biochemical cascade involves complex parallel recruitment pathways of certain SAC proteins (like BUBR1). 
% the aforementioned difficulty caused by 
% The formation of the CDC20-MAD2 subcomplex is rate-limiting, as it involves an induced conformational change of MAD2.
% The revelation of how different molecules orchestrate to build a de facto logic gate from biochemical reactions in the SAC will further our appreciation of evolution,
%  Then, we looked into whether auxiliary motifs on KNL1 that may